---
title: "Replicate paper draft"
author: "Xiaona Zhou"
date: "6/23/2020"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  pdf_document:
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
library(scales)
library(tidyverse)
library(plotly)
library(lubridate)
library(broom)
library(modelr)
library(readxl)
library(zoo)
library(caret)
theme_set(theme_bw())

knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

# Introduction
We are replicating the result from paper [Predicting the Present with Google Trends.](http://people.ischool.berkeley.edu/~hal/Papers/2011/ptp.pdf)

# Replicate results from 3.1 Motor vehicles and parts
## Replication with authors' data
### Replicate baseline seasonal AR-1 model

We began the replcation process by using the data that was used for the paper. According to the paper, the model follows the form of $y_t = b_1y_{t-1}+b12y_{t-12}+e_t$(AR-1 model). We used the codes below to obtain the same summary table to the model.

```{r}
# load the data that was used for the paper
merged <- read_csv("merged.csv")

# take the log of all sales values
merged$sales<-log(merged$sales)

# apply lm(). lag() is used to capture y_t-1 and y_t-12
model1 <- lm(data = merged, sales~lag(sales, 1)+lag(sales,12))

#the summary of the model
summary(model1)
```

The summary we obtained is identical to the paper. 


### Replicate the model with trends

According to the paper, we added two trends, *Trucks & SUVs* and *Automotive Insurance* to the previous model. 

```{r}
# model with trends
model_with_trend <- lm(data = merged, sales~lag(sales, 1)+lag(sales,12) + suvs + insurance)

# summary table of the model
summary(model_with_trend)
```

Once again, The summary we obtained is identical to the paper.

### Replicate figure 2 

The authors used rolling window forecast to confirm that the Trends variables improve out-of-sample forecasting. 
What is rolling window forecast? 
That's say now is July, 2015. We would use all the actual data that is available to train a model and predict August, 2015. We'll keep this predicted value aside. Imagine now is August, 2015, we have all the atual data(up to July, 2015) avaiable to train another model to predict September, 2015. Repeat this process for baseline model and the model with trends. 

```{r}

# creating base 
base <- merged # a place holder for predicted sales from baseline model
# rolling window forcast
for (i in 18:91){
  merged_t <- merged[1:i-1,]
  model1 <- lm(data = merged_t, sales~lag(sales, 1)+lag(sales,12))
  base$sales[i] <- predict(model1,merged[1:i,])[i]
}
base <- base[18:91,]# select out all the predicted sales


# creating trends 
trends <- merged
for (i in 18:91){
  merged_t <- merged[1:i-1,]
  model1 <- lm(data = merged_t, sales~lag(sales, 1)+lag(sales,12)+ suvs + insurance)
  trends$sales[i] <- predict(model1,merged[1:i,])[i]
}
trends <- trends[18:91,]


# Make the graph
# add labels to the data
actual <- merged[18:91,] %>% 
  mutate(label ="actual")
base <- base %>% 
  mutate(label = "base")
trends <- trends %>% 
  mutate(label ="trends")

# means absolute error
mean(abs(trends$sales-actual$sales))
mean(abs(base$sales-actual$sales))

# data for the recession period Dec 2007 to June 2009
recession_trends <- trends %>% 
  filter(Period>="2007-12-01"& Period<="2009-06-01")
recession_base <- base %>% 
  filter(Period>="2007-12-01"& Period<="2009-06-01")
recession_actual <- actual %>% 
  filter(Period>="2007-12-01"& Period<="2009-06-01")

# means absolute error for the recession period
# MAE for the model with trends
mean(abs(recession_trends$sales-recession_actual$sales))

# MAE for the model for baseline model
mean(abs(recession_base$sales-recession_actual$sales))


# Overall improvement 
(mean(abs(base$sales-actual$sales))-mean(abs(trends$sales-actual$sales)))/mean(abs(base$sales-actual$sales))


# recession improvement
(mean(abs(recession_base$sales-recession_actual$sales))-mean(abs(recession_trends$sales-recession_actual$sales)))/mean(abs(recession_base$sales-recession_actual$sales))

# Note: the improvements they stated in the paragraph were different from what they labeled on the graph. First, I thought they used some kind of function, but the MAE() from library(caret) gives same value.


# R^2 for base
(cor(base$sales,actual$sales))^2

# R^2 for trends
(cor(trends$sales,actual$sales))^2

# combine data for plotting
plot_data <- rbind(actual, base, trends)
ggplot(plot_data, aes(x=Period, y = sales, color = label, linetype = label))+
  geom_line()+
  scale_colour_manual(values=c("black", "red","grey"))+
  scale_linetype_manual(values = c("solid", "dashed", "solid"))+
  ylab('log(mvp)')+
  xlab('Index')+
  labs(subtitle ="MAE improvement: Overall = 10.66%  During recession = 21.46%" )
```


## Replication with raw data

### Where did we find the data?

*The sales data*:
We first tried the [link](https://www.census.gov/retail/marts/www/timeseries.html) provided in the paper, but the data is adjusted. We were able to find the unadjusted data through this [link](https://www.census.gov/econ/currentdata/dbsearch?program=MARTS&startYear=2004&endYear=2020&categories%5B%5D=441&dataType=SM&geoLevel=US&notAdjusted=1&submit=GET+DATA&releaseScheduleId=). However, the values were different from the original data(used by the authors of the paper). The difference increases as year increases. The reason for that is still unclear. We will compare the two data set in the next session.

*The search data*:
[Trucks & SUVs](https://trends.google.com/trends/explore?cat=610&date=2004-01-01%202011-07-01&geo=US)
[Auto Insurance](https://trends.google.com/trends/explore?cat=467&date=2004-01-01%202011-07-01&geo=US)


### Comparing sales data used in the paper with data we obtained from census.gov
We noticed that the sales from the data we obtained were always less than the data used in the paper. We want to find out why.

```{r}
# load the data
sales <- read_excel("sales.xls")
merged <- read_csv("merged.csv")
# Period has type char in the data, convert that to yearmonth 
ym <- as.yearmon(sales$Period, "%b-%Y")
# use as.Date() to convert the type of ym to date
sales$Period <- as.Date(ym)
#keep only data from 01/2004 through 07/2011 and renames the data frame as unadj, rename the column Value as sales,and add label "unadjusted"
unadj<- sales %>% 
  filter(Period <="2011-07-01") %>% 
  mutate(label = "unadjusted") %>% 
  rename(sales = Value)
# convert sales to numeric 
unadj$sales <-as.numeric(unadj$sales)
# add label to original data, keep only Period, sales and label
orig <- merged %>% 
  mutate(label = "original") %>% 
  select(-insurance, -suvs)
# stack two data frame
com_data <- rbind(unadj, orig)

#plot log(sales) Vs Period 
ggplot(com_data, aes(x=Period, y = log(sales), color = label)) +
  geom_line()

# plot orignal against unadjusted
joined_data <- unadj %>% select(-label) %>% rename(unadjusted_sales = sales) %>% left_join(orig, by ="Period") %>% select(-label)

ggplot(joined_data, aes(x = sales, y = unadjusted_sales)) +
  geom_point()+
  geom_abline(linetype = "dashed") +
  xlab('original') +
  ylab('unadjusted')+
  ggtitle("original sales against unadjusted sales")
```

From the first graph, we can see that the sales values from both original and unajusted are very close to each other(almost overlapping). When I plotted originaldata against unadjusted data, they almost lined up on the diagonal. Note that original is always less than unadjusted. I wonder if people went back and modified that data after the authors obtained the data. Further investigation is needed. In conclusion, the unadjusted data we found is very close to the original data. 

### Attempts to find the transformation on Google Trends data in original data

The authors did not specify what kind of transformation/normalization/standarlization they performed on Google Trends data. 
Below is what we have tried to transform trends data.
1. new_value = (old_value-first_observed_value)/max(old_value)
2. Min-Max Normalization
3. Use scale()

None of these match with the original data. Therefore, we decided to process without any transformtion on Google Trends data. 

We also tried other key words like "Buy car", "car dealer", "Vehicle shop", but none of these were significant to the model

### Replicate the results again with unadjusted data from census.gov with trends data from original paper

We would like to compare the outputs of unadjusted data with the outputs from the original data.

```{r}
#take the log of sales, take out the label column
unadj_rep <- unadj %>% mutate(sales=log(sales)) %>% select(-label)

# google_trends contains trends data and Period for join
google_trends <- merged %>% select(-sales)

# join trends to unadj_rep
unadj_with_trends <- unadj_rep %>% left_join(google_trends, by = "Period")

# replicate baseline model
model1 <- lm(data = unadj_with_trends, sales~lag(sales, 1)+lag(sales,12))

#the summary of the model
summary(model1)


# replicate the model with trends
model_with_trend <- lm(data = unadj_with_trends, sales~lag(sales, 1)+lag(sales,12) + suvs + insurance)
summary(model_with_trend)



# replicate figure 2

# creating base 
base_unajusted <- unadj_with_trends
for (i in 18:91){
  merged_t <- unadj_with_trends[1:i-1,]
  model1 <- lm(data = merged_t, sales~lag(sales, 1)+lag(sales,12))
  base_unajusted$sales[i] <- predict(model1,unadj_with_trends[1:i,])[i]
}
base_unajusted <- base_unajusted[18:91,]


# creating trends 
trends_unajusted <- unadj_with_trends
for (i in 18:91){
  merged_t <- unadj_with_trends[1:i-1,]
  model1 <- lm(data = merged_t, sales~lag(sales, 1)+lag(sales,12) + suvs + insurance)
  trends_unajusted$sales[i] <- predict(model1,unadj_with_trends[1:i,])[i]
}
trends_unajusted <- trends_unajusted[18:91,]


# label different data sets
actual_unajusted <- unadj_with_trends[18:91,]
actual_unajusted <- actual_unajusted %>%
  mutate(label ="actual_unajusted")
base_unajusted <- base_unajusted %>%
  mutate(label = "base_unajusted")
trends_unajusted <- trends_unajusted %>%
  mutate(label ="trends_unajusted")

# R^2 for base
(cor(base_unajusted$sales,actual_unajusted$sales))^2

# R^2 for trends
(cor(trends_unajusted$sales,actual_unajusted$sales))^2


# stack all data sets for plotting
plot_data <- rbind(actual, base, trends, actual_unajusted, base_unajusted, trends_unajusted)
ggplot(plot_data, aes(x=Period, y = sales, color = label, linetype = label))+
  geom_line()+
  scale_colour_manual(values=c("black","black", "red","red","grey","grey"))+
  scale_linetype_manual(values = c("solid","solid","dashed", "dashed", "solid","solid"))+
  ylab('log(mvp)')+
  xlab('Index')
```

From the graph, we see that lines are almost overlaped, and $R^2$ are very close to the ones from original data. 




### Replicate the results with unadjusted data from census.gov and search data from Google Trends without unknown transformations 


```{r}
# trend 1 
# search data in category Trucks & SUVs between 01/01/2004 and 07/01/2011
suvs_trends <- read_csv("New_Trucks_suvs.csv")
names1<- names(suvs_trends)
suvs_trends <- suvs_trends %>%
  rename(suvs = names1[2],
         Period = Month)

# trend 2
# search data in category Auto Insurance between 01/01/2004 and 07/01/2011
insurance_trends <- read_csv("New_auto_insurance.csv")
names2 <- names(insurance_trends)
insurance_trends <- insurance_trends %>%
  rename(insurance = names2[2],
         Period = Month)

# census data 
unadj_full <- sales %>% filter(Period <= "2011-07-01")

#join trends data
trends_full <- left_join(insurance_trends, suvs_trends, by = "Period") %>% mutate(Period = as.Date(as.yearmon(Period, "%Y-%m")))

# join all data
with_trends_full <- left_join(unadj_full, trends_full, by = "Period") %>% rename(sales = Value) 

# take the log of sales
with_trends_full$sales = log(as.numeric(with_trends_full$sales))


# apply lm(). lag() is used to capture y_t-1 and y_t-12
model1 <- lm(data = with_trends_full, sales~lag(sales, 1)+lag(sales,12))


# summary of model1
tidy(model1)
m1 <- glance(model1)
# model with trends
model_with_trend <- lm(data = with_trends_full, sales~lag(sales, 1)+lag(sales,12) + insurance+ suvs)

# summary table of the model
# summary(model_with_trend)
tidy(model_with_trend)
mt<-glance(model_with_trend)
```

The $R^2$ for base model is `r m1$r.squared`, and The $R^2$ for base model is `r mt$r.squared`. We see an improvement in overall fitting. We will further investigate if this is true with out-of-sample forcasting.


### Evaluate the model by using rolling window forecast starting from 2005-07-01

```{r}
base <- with_trends_full
# creating predicted base with rolling window nowcast
for (i in 18:91){
  merged_t <- with_trends_full[1:i-1,]
  model1 <- lm(data = merged_t, sales~lag(sales, 1)+lag(sales,12))
  base$sales[i] <- predict(model1,with_trends_full[1:i,])[i]
}
base <- base[18:91,]


# creating predicted trends with rolling window nowcast

trends <- with_trends_full
for (i in 18:91){
  merged_t <- with_trends_full[1:i-1,]
  model1 <- lm(data = merged_t, sales~lag(sales, 1)+lag(sales,12)+ suvs + insurance)
  trends$sales[i] <- predict(model1,with_trends_full[1:i,])[i]
}
trends <- trends[18:91,]

# Make the graph

actual <- with_trends_full[18:91,]
actual <- actual %>% 
  mutate(label ="actual")
base <- base %>% 
  mutate(label = "base")
trends <- trends %>% 
  mutate(label ="trends")
plot_data <- rbind(actual, base, trends) 
ggplotly(ggplot(plot_data, aes(x=Period, y = sales, color = label, linetype = label))+
  geom_line()+
  scale_colour_manual(values=c("black", "red","grey"))+
  scale_linetype_manual(values = c("solid", "dashed", "solid"))+
  ylab('log(mvp)')+
  xlab('Index'))
```


##### Calculate Mean Absolute Error and R^2 (with raw data)

```{r}
# means absolute error
MAE_base <- mean(abs(base$sales-actual$sales))
MAE_trends <- mean(abs(trends$sales-actual$sales))

# data for recession period Dec 2007 to June 2009
recession_trends <- trends %>% 
  filter(Period>="2007-12-01"& Period<="2009-06-01")
recession_base <- base %>% 
  filter(Period>="2007-12-01"& Period<="2009-06-01")
recession_actual <- actual %>% 
  filter(Period>="2007-12-01"& Period<="2009-06-01")
MAE_recession_trends <- mean(abs(recession_trends$sales-recession_actual$sales))
MAE_recession_base <- mean(abs(recession_base$sales-recession_actual$sales))


# Overall improvement 
overall_improv <- (mean(abs(base$sales-actual$sales))-mean(abs(trends$sales-actual$sales)))/mean(abs(base$sales-actual$sales))


# recession improvement
recession_improv <- (mean(abs(recession_base$sales-recession_actual$sales))-mean(abs(recession_trends$sales-recession_actual$sales)))/mean(abs(recession_base$sales-recession_actual$sales))



#R^2 for base
r_sqr_base <-(cor(base$sales,actual$sales))^2

# R^2 for trends
r_sqr_trends <- (cor(trends$sales[3:73],actual$sales[3:73]))^2

#We can use MAE() from caret package to calculate MAE
# 
# library(caret)
#  
# # R native funcitons
# MAE(recession_trends$sales, recession_actual$sales)
# MAE(recession_base$sales, recession_actual$sales)
# 
# MAE(trends$sales, actual$sales)
# MAE(base$sales, actual$sales)

```

The MAE for base model is `r MAE_base`, and the MAE for the model with trends is `r MAE_trends`. The overall improvement is `r overall_improv*100`%. That means adding trends to the model make things worse. We see a totally different story when forces on recession period. The MAE for base model is `r MAE_recession_base`, and the MAE for the model with trends is `r MAE_recession_trends` during the recession. There is `r recession_improv*100`% improvement after adding trends to the model. 



# Extensions

### Replication using all available data  01/01/2004 -05/01/2020 

```{r}
suvs_trends <- read_csv("trucks_suvs_2004_2020.csv")
insurance_trends <- read_csv("auto_insurance_2004_2020.csv")
# rename
names1<- names(suvs_trends)
names2 <- names(insurance_trends)
suvs_trends <- suvs_trends %>% 
  rename(suvs = names1[2],
         Period = Month)
insurance_trends <- insurance_trends %>% 
  rename(insurance = names2[2], 
         Period = Month)

unadj_full <- sales %>% filter(Period <= "2020-05-01")

trends_full <- left_join(insurance_trends, suvs_trends, by = "Period") %>% mutate(Period = as.Date(as.yearmon(Period, "%Y-%m"))) 
with_trends_full <- left_join(unadj_full, trends_full, by = "Period") %>% rename(sales = Value) 
with_trends_full$sales = log(as.numeric(with_trends_full$sales))

# apply lm(). lag() is used to capture y_t-1 and y_t-12
model1 <- lm(data = with_trends_full, sales~lag(sales, 1)+lag(sales,12))

#the summary of the model
summary(model1)

# model with trends
model_with_trend <- lm(data = with_trends_full, sales~lag(sales, 1)+lag(sales,12) + suvs + insurance)

# summary table of the model
summary(model_with_trend)

# Rolling window 
base <- with_trends_full
len <- nrow(base)
# creating predicted base with rolling window forecast
for (i in 18:len){
  merged_t <- with_trends_full[1:i-1,]
  model1 <- lm(data = merged_t, sales~lag(sales, 1)+lag(sales,12))
  base$sales[i] <- predict(model1,with_trends_full[1:i,])[i]
}
base <- base[18:len,]


# creating predicted trends with rolling window forecast
trends <- with_trends_full
for (i in 18:len){
  merged_t <- with_trends_full[1:i-1,]
  model1 <- lm(data = merged_t, sales~lag(sales, 1)+lag(sales,12)+ suvs + insurance)
  trends$sales[i] <- predict(model1,with_trends_full[1:i,])[i]
}
trends <- trends[18:len,]

# Make the graph
actual <- with_trends_full[18:len,]
actual <- actual %>% 
  mutate(label ="actual")
base <- base %>% 
  mutate(label = "base")
trends <- trends %>% 
  mutate(label ="trends")
plot_data <- rbind(actual, base, trends)
ggplotly(ggplot(plot_data, aes(x=Period, y = sales, color = label, linetype = label))+
  geom_line()+
  scale_colour_manual(values=c("black", "red","grey"))+
  scale_linetype_manual(values = c("solid", "dashed", "solid"))+
  ylab('log(mvp)')+
  xlab('Index'))
```

*Add commends here* 


## Forecasting with new model

We want forcast by using the model $y_t+2=b_1y_{t}+b_2y_{t-12} + e_t$. We'll use rolling window forcast to evaluate our model. At the end, We will try to predict the sales of July, 2020. 

commend on *yearly trends*

```{r}

## getting data ready
suvs_trends <- read_csv("trucks_suvs_2004_2020.csv")
insurance_trends <- read_csv("auto_insurance_2004_2020.csv")
# rename
names1<- names(suvs_trends)
names2 <- names(insurance_trends)
suvs_trends <- suvs_trends %>% 
  rename(suvs = names1[2],
         Period = Month)

insurance_trends <- insurance_trends %>% 
  rename(insurance = names2[2], 
         Period = Month)

unadj_full <- sales %>% filter(Period <= "2020-05-01")

trends_full <- left_join(insurance_trends, suvs_trends, by = "Period") %>% mutate(Period = as.Date(as.yearmon(Period, "%Y-%m"))) 
with_trends_full <- left_join(unadj_full, trends_full, by = "Period") %>% rename(sales = Value) 
with_trends_full$sales = log(as.numeric(with_trends_full$sales))
#with_trends_full$sales = as.numeric(with_trends_full$sales)


## starting from 2018
whole_data_2018 <- with_trends_full %>% 
  filter(Period >="2015-01-01") # we want to predict 2018-01-01 when it is 2017-10-01, we need data from 2016-10-01 
#months <- nrow(whole_data_2018)
months <- 60
 
base <- whole_data_2018
start <- 37
# creating predicted base with rolling window nowcast
for (i in start:months){
  merged_t <- whole_data_2018[1:i-1,]
  model1 <- lm(data = merged_t, sales~lag(sales, 2)+lag(sales,14))
  base$sales[i] <- predict(model1,whole_data_2018[1:i,])[i]
}
base <- base[start:months,]


# creating predicted trends with rolling window nowcast

trends <- whole_data_2018
for (i in start:months){
  merged_t <- whole_data_2018[1:i-1,]
  model1 <- lm(data = merged_t, sales~lag(sales, 2)+lag(sales,14)+ suvs + insurance)
  trends$sales[i] <- predict(model1,whole_data_2018[1:i,])[i]
}
trends <- trends[start:months,]

# Make the graph

actual <- whole_data_2018[start:months,]
actual <- actual %>% 
  mutate(label ="actual")
base <- base %>% 
  mutate(label = "base")
trends <- trends %>% 
  mutate(label ="trends")
plot_data <- rbind(actual, base, trends) 
ggplotly(ggplot(plot_data, aes(x=Period, y = sales, color = label, linetype = label))+
  geom_line()+
  scale_colour_manual(values=c("black", "red","grey"))+
  scale_linetype_manual(values = c("solid", "dashed", "solid"))+
  ylab('log(mvp)')+
  xlab('Index'))



MAE(trends$sales, actual$sales)
MAE(base$sales, actual$sales)
(MAE(base$sales, actual$sales)-MAE(trends$sales, actual$sales))/MAE(base$sales, actual$sales)
#R^2 for base
(cor(base$sales,actual$sales))^2

# R^2 for trends
(cor(trends$sales,actual$sales))^2
```


The model did very bad. observed yearly trends

```{r}
# yearly trends
ggplotly(ggplot(with_trends_full, aes(x = Period, y = sales)) + geom_line())
```



 
with model $y_t = b_1y_{t-1}+b12y_{t-12}+e_t$(AR-1 model) much better performance
```{r}

## getting data ready
suvs_trends <- read_csv("trucks_suvs_2004_2020.csv")
insurance_trends <- read_csv("auto_insurance_2004_2020.csv")
# rename
names1<- names(suvs_trends)
names2 <- names(insurance_trends)
suvs_trends <- suvs_trends %>% 
  rename(suvs = names1[2],
         Period = Month)

insurance_trends <- insurance_trends %>% 
  rename(insurance = names2[2], 
         Period = Month)

unadj_full <- sales %>% filter(Period <= "2020-05-01")

trends_full <- left_join(insurance_trends, suvs_trends, by = "Period") %>% mutate(Period = as.Date(as.yearmon(Period, "%Y-%m"))) 
with_trends_full <- left_join(unadj_full, trends_full, by = "Period") %>% rename(sales = Value) 
with_trends_full$sales = log(as.numeric(with_trends_full$sales))
#with_trends_full$sales = as.numeric(with_trends_full$sales)

ggplotly(ggplot(with_trends_full, aes(x = Period, y = sales)) + geom_line())

## starting from 2018
whole_data_2018 <- with_trends_full %>% 
  filter(Period >="2015-01-01") # we want to predict 2018-01-01 when it is 2017-10-01, we need data from 2016-10-01 
#months <- nrow(whole_data_2018)
months <- 60
 
base <- whole_data_2018
start <- 37
# creating predicted base with rolling window nowcast
for (i in start:months){
  merged_t <- whole_data_2018[1:i-1,]
  model1 <- lm(data = merged_t, sales~lag(sales, 1)+lag(sales,12))
  base$sales[i] <- predict(model1,whole_data_2018[1:i,])[i]
}
base <- base[start:months,]


# creating predicted trends with rolling window nowcast

trends <- whole_data_2018
for (i in start:months){
  merged_t <- whole_data_2018[1:i-1,]
  model1 <- lm(data = merged_t, sales~lag(sales, 1)+lag(sales,12)+ suvs + insurance)
  trends$sales[i] <- predict(model1,whole_data_2018[1:i,])[i]
}
trends <- trends[start:months,]

# Make the graph

actual <- whole_data_2018[start:months,]
actual <- actual %>% 
  mutate(label ="actual")
base <- base %>% 
  mutate(label = "base")
trends <- trends %>% 
  mutate(label ="trends")
plot_data <- rbind(actual, base, trends) 
ggplotly(ggplot(plot_data, aes(x=Period, y = sales, color = label, linetype = label))+
  geom_line()+
  scale_colour_manual(values=c("black", "red","grey"))+
  scale_linetype_manual(values = c("solid", "dashed", "solid"))+
  ylab('log(mvp)')+
  xlab('Index'))



MAE(trends$sales, actual$sales)
MAE(base$sales, actual$sales)
(MAE(base$sales, actual$sales)-MAE(trends$sales, actual$sales))/MAE(base$sales, actual$sales)
#R^2 for base
(cor(base$sales,actual$sales))^2

# R^2 for trends
(cor(trends$sales,actual$sales))^2
```



## Predict June and July 2020 with variouse models 

1. With model $y_t = b_1y_{t-1}+b12y_{t-12}+e_t$(AR-1 model)

```{r}

data_for_pre_20 <- with_trends_full %>% 
  filter(Period >="2017-10-01"& Period <="2019-12-01")
model1 <- lm(data = data_for_pre_20 , sales~lag(sales, 1)+lag(sales,12))
data_2020 <- with_trends_full %>% filter(Period >="2019-01-01")
summary(model1)

predicted <- predict(model1,data_2020)[13:17]
MSE <- mean((predicted-data_2020$sales[13:17])^2)
mse(model1, data_2020)


#We will predict June 2020 and then July 2020

n <- tidy(model1)

June <- exp(n$estimate[1] +n$estimate[2]*data_2020$sales[17]+n$estimate[3]*data_2020$sales[6])
June

July <-  exp(n$estimate[1] +n$estimate[2]*log(June)+n$estimate[3]*data_2020$sales[7])

July
```


2. With model $y_t = b_1y_{t-1}+b_2y_{t-12}+e_t+b_3insurance+b_4SUVs$
```{r}
data_for_pre_20 <- with_trends_full %>% 
  filter(Period >="2017-10-01"& Period <="2019-12-01")
model_with_trend <- lm(data = data_for_pre_20 , sales~lag(sales, 1)+lag(sales,12) + insurance + suvs)
data_2020 <- with_trends_full %>% filter(Period >="2019-01-01")
summary(model_with_trend )

predicted <- predict(model_with_trend ,data_2020)[13:17]
MSE <- mean((predicted-data_2020$sales[13:17])^2)
mse(model_with_trend , data_2020)


#We will predict July 2020

n <- tidy(model_with_trend )

June <- exp(n$estimate[1] +n$estimate[2]*data_2020$sales[17]+n$estimate[3]*data_2020$sales[6]+n$estimate[4]*data_2020$insurance[17]+n$estimate[5]*data_2020$suvs[17])
June
```

2. With model $y_t = b_1y_{t-2}+b_2y_{t-14}+e_t+b_3insurance+b_4SUVs$(AR-1 model) *Bad model*


```{r}
data_for_pre_20 <- with_trends_full %>% 
  filter(Period >="2017-10-01"& Period <="2019-12-01")
model_with_trend <- lm(data = data_for_pre_20 , sales~lag(sales, 2)+lag(sales,14) + insurance + suvs)
data_2020 <- with_trends_full %>% filter(Period >="2019-01-01")
summary(model_with_trend )

predicted <- predict(model_with_trend ,data_2020)[13:17]
MSE <- mean((predicted-data_2020$sales[13:17])^2)
mse(model_with_trend , data_2020)


#We will predict July 2020

n <- tidy(model_with_trend )

July <- exp(n$estimate[1] +n$estimate[2]*data_2020$sales[17]+n$estimate[3]*data_2020$sales[5]+n$estimate[4]*data_2020$insurance[17]+n$estimate[5]*data_2020$suvs[17])
July-115180.4
```



3. $y_t=b_1y_{t-12}+b_2y_{t-24}+b_3y_{t-36}+b_4y_{t-48}+b_5insurance + b_6SUVs$

```{r}
data_for_pre_20 <- with_trends_full %>% 
  filter(Period >="2010-01-01"& Period <="2016-12-01")
# model_with_trend <- lm(data = data_for_pre_20 , sales~lag(sales, 12)+lag(sales, 24)+lag(sales, 36)+lag(sales, 48)+ insurance + suvs)
model_with_trend <- lm(data = data_for_pre_20 , sales~lag(sales, 12)+lag(sales, 48)+ insurance + suvs)

data_2020 <- with_trends_full %>% filter(Period >="2013-01-01")
summary(model_with_trend )

predicted <- predict(model_with_trend ,data_2020)[49:89]
MSE <- mean((predicted-data_2020$sales[49:89])^2)
mse(model_with_trend , data_2020)


#We will predict July 2020

n <- tidy(model_with_trend )

July <- exp(n$estimate[1] +n$estimate[2]*data_2020$sales[17]+n$estimate[3]*data_2020$sales[5]+n$estimate[4]*data_2020$insurance[17]+n$estimate[5]*data_2020$suvs[17])
July-115180.4
```











